Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-02-02 grove-paper.tex
updates from call w/ Cyrus

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \#u \\
      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \#u \\
      e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \#u \\
    \end{array}
  \]
  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
  \label{fig:Base syntax}
\end{figure}

@@

% TODO add references into multiparent store and cycles store

@@

A grove is a 4-tuple of the following form:
\[
  (e_r, MP, NP, W)
\]
where
\begin{align*}
  e_r &= \expr(v_0) \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
\end{align*}

The first component $e_r$ is the expression rooted at the root vertex $v_0$.
The second component $MP$ is the set of expressions rooted at a vertex with multiple parents.
The third component $NP$ is the set of expressions rooted at a vertex with no parents.
The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent. In each of these, $\parents(v)$ is defined as follows.  \[ \parents(v) = \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\}  \] \\

For example, the grove corresponding to the graph in Figure~XXX is ...

@@

Figure YYY defines $\expr(v)$ wich determines the expression for a given vertex $v$.
There's one rule for each constructor form, e.g., $\ExpApp$.
In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).

@@

\begin{figure*}
\begin{align*}
  e_r &= \expr(v_0) \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
\end{align*}

\begin{align*}
  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
\end{align*}

\begin{displaymath}
\begin{array}{l}
  \expr(v=(u, Root)) = expr'(v, Root) \\
  \expr(v=(u, \ExpVar(x))) = x^u \\
  \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
  \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
\end{array}
\end{displaymath}

@@

\begin{align*}
  \expr'(v,p) = \begin{cases}
    \hole & \children(v,p) = \varnothing \\
    \expr(v') & \children(v,p) = \{v'\} \\
    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| > 1 \\
    \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \land v' = \min(\ancestors(v')) \\
    \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
  \end{cases}
\end{align*}

>>>>

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \multiVertex{u} \mid \cycleVertex{u} \\
      e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \multiVertex{u} \mid \cycleVertex{u} \\
    \end{array}
  \]
  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
  \label{fig:Base syntax}
\end{figure}

@@

@@

@@

% \newcommand{\multiVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\curlyveedownarrow_{#1}}}}}
% \newcommand{\cycleVertex}[1]{\textcolor{red}{\ensuremath{\mathbf{\rcirclearrowleft_{#1}}}}}

A grove is a 4-tuple of the following form:
\[
  (e_r, MP, NP, W)
\]
where
\begin{align*}
  e_r &= \expr(v_0) \\
  MP &= \{\expr(v) : |\parents(v)| > 1\} \\
  NP &= \{\expr(v) : |\parents(v)| = 0 \land v \ne v_0\} \\
  W  &= \{\expr(v) : |\parents(v)| = 1 \wedge v = \min(\ancestors(v))\}
\end{align*}
For example, the grove corresponding to the graph in \autoref{fig:Decomposition example} is
\begin{align*}
  e_r &= \hole~\texttt{*}^{46}~\#48~\texttt{+}^{44}~\#48 \\
  MP &= \{x^{48}\} \\
  NP &= \{z^{58}\} \\
  W &= \{\hole~\texttt{+}^{52}~\#52~\texttt{*}^{54}~y^{56}\} \\
\end{align*}

Let us now examine each of the four components in more detail.
The first component $e_r$ is the expression derived from the root vertex $v_0$.
\autoref{fig:Decomposition expr} defines $\expr(v)$ wich determines the expression derived from $v$.
There's one rule for each constructor form, e.g., $\ExpApp$.
In each child position, e.g., $\AppFun, \AppArg$, we call an auxiliary function $\expr'(v, p)$ which ... (explain different cases and what's going on here).

The second component $MP$ is the set of expressions derived from a vertex with multiple parents.
Wherever an edge to such a vertex appears, we place a reference to it $\#u$.
In the example, $x^{48}$ is referenced twice in $e_r$.

The third component $NP$ is the set of expressions rooted at a vertex with no parents, other than the root vertex.
In the example, $z^{58}$ has no parents (because it was deleted), so it appears only in $NP$.

The fourth component $W$ is the set of expressions derived from the remaining vertices, which must necessarily form cyclic structures we call wreaths, where each vertex has exactly one parent.

In $W$, $\ancestors(v)$ is the set of vertices for which there is a path to $v$.
Note that $v = \min(\ancestors(v))$ only if $v$ is part of a cycle.

% TODO: replace # with \multiVertex and \cycleVertex
% TODO: update \multiChild with more than 2 children and use it for conflict holes
% TODO: gray out ids in the grammar and expressions
% TODO: add a child to z^58 and rotate 51-52
% TODO: move example grove to right of its diagram
% TODO: distinguish between multi-parent and cycle refs

\begin{figure}
  \begin{displaymath}
    \begin{array}{l}
      \expr(v=(u, Root)) = expr'(v, Root) \\
      \expr(v=(u, \ExpVar(x))) = x^u \\
      \expr(v=(u, \ExpApp)) = (\expr'(v, \AppFun)~\expr'(v, \AppArg))^u \\
      \expr(v=(u, \ExpLam)) = \lambda^u \patt'(v, \LamParam) : \type'(v, \LamType) . \expr'(v, \LamBody) \\
    \end{array}
  \end{displaymath}

  \begin{align*}
    \expr'(v,p) = \begin{cases}
      \hole & \children(v,p) = \varnothing \\
      \expr(v') & \children(v,p) = \{v'\} \land |\parents(v')| = 1 \\
                & \phantom{\children(v,p)} \land v' \ne \min(\ancestors(v')) \\
      \#u & \children(v,p) = \{v'\} \land |\parents(v')| > 1 \\
      \#u & \children(v,p) = \{v'=(u,k)\} \land |\parents(v')| = 1 \\
          & \phantom{\children(v,p)} \land v' = \min(\ancestors(v')) \\
      \{\expr(v_1) | \ldots | \expr(v_n)\} & \children(v,p) = \{v_1, \ldots, v_n\} \\
    \end{cases}
  \end{align*}

  \begin{align*}
    \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
    \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v)) \\
    \children(v, p) &= \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\} \\
    \min(\{v_1 = (u_1, k_1), \ldots, v_n = (u_n, k_n)\}) &= (u_j, k_j) \text{ s.t. } j \in \{1, \ldots, n\} \land u_j \le u_i \forall i \in \{1, \ldots, n\}
      \end{align*}
  \caption{Caption2 goes here. Note that $\ancestors(v)$ is defined as the least fixed point of the equation above.}
  \label{fig:Decomposition expr}
\end{figure}

@@

\begin{figure}[b]
\begin{align*}
  \parents(v) &= \{v' : \exists \e = (u, v', p, v), G(\e) = \Plus\} \\
  \ancestors(v) &= \parents(v) \cup \ancestors(\parents(v))
  \children(v, p) = \{v' : \exists \e (u, v, p, v'), G(\e) = \Plus\}
\end{align*}

@@
